Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2017-08-30 stp/paper_aaai/main.tex
Added minizinc model at the end, shortened text. added description of serch startegy

\usepackage{color,colortbl}
\usepackage{xcolor}
\definecolor{Gray}{gray}{0.96}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

@@

\usepackage{supertabular}

@@

\subsubsection{Example} Consider the situation where we have some simple relational database tables with columns $\langle$\textit{Surname, Workplace, Date, Location}$\rangle$, 
$\langle$\textit{Company, Event, Begin, Until, Location}$\rangle$ and 
$\langle$\textit{Name, D.O.B, Location}$\rangle$.
We do not \textit{a priori} know whether ``Date'' in the first table is the 
date of birth or the date a persons started working; or whether the 
``Location'' on the second table refers to the location of the event, or the 
address of the company. 
If we are given an ontology like the one in Fig. \ref{FIG:onto}, we would like to automatically map the new data sources to classes of the 
ontology.\hfill$\square$

@@

In our work we consider that an ontology $\mathcal{O}$ includes basic elements 
as classes which represent a set of concepts, literal values, individuals which 
are members of classes and properties~\cite{Spanos:semweb}.
Properties are classified into \emph{object properties}, which relate two 
individuals, 
and \emph{datatype properties}, which relate individuals to literal values.
An example of an ontology is given in Fig.~\ref{FIG:onto}. 
Here, an individual of the class ``Organisation'' can be related to an individual of the class ``Person'' via object properties ``ceo'' or ``worksFor'', and it can have a data property ``name''. 
We also consider a special ``subclass'' (or ``is-a'') type of object properties.
For 
example, ``Cities'' and ``States'' are both ``Places''. 
A subclass inherits all the properties of the parent class.

@@

The algorithm we use for the construction of the alignment graph 
$\mathcal{G_O}$ was given by~\cite{taheriyan2016learning}.
Briefly, the algorithm has three parts:
%\begin{enumerate}
$(i)$ %\item 
Adding the known semantic models
$(ii)$ %\item 
Adding the semantic types learned for the target source
%\item 
$(iii)$ Expanding the graph using the domain ontology $\mathcal{O}$.
%\end{enumerate}

@@

The graph is weighted by a function $w_\mathcal{O} : \mathcal{E_O} \mapsto 
\mathbb{R}$ in 
such a way that edges which are present in
the known semantic models have lower weights than those 
which are inferred from the ontology.
As we will see in the next section, this makes edges present in semantic models 
more attractive. 
Details on the weighting function can be found in \cite{taheriyan2016learning}.
%\ddg{How exactly are they weighted?}

@@

A Constraint Optimization Problem (COP) is a tuple $P = (\boldsymbol{v},C,o)$ 
where $\boldsymbol{v}$ is a 
set of \emph{variables}, $C$ is a set of $n$-ary 
\emph{constraints} over variables $\boldsymbol{v}$ and $o$ is an 
\emph{objective function} $o : \boldsymbol{v} \mapsto \mathbb{R}$ to be 
(w.l.o.g.) minimized. 
A valuation such that all variables in $\boldsymbol{v}$ map 
to exactly one value, and all the constraints in $C$ are satisfied by the 
valuation is a \emph{solution} to $P$. A solution $\theta$ is optimal if$\not 
\exists \theta', \theta'(\boldsymbol{v}) < \theta(\boldsymbol{v})$.

@@

Briefly, the way a CP solver works is by assigning a value to each variable of 
$\boldsymbol{v}$ in turn. At each iteration it will check whether any 
constraint in $C$ is violated, in which case it will backtrack to change its 
last decision. \ignore{The choice of the order of the variables to be assigned, 
and the 
choice of the value to be assigned to each variable is called the \emph{search 
strategy}.}
CP uses a combination of search and \emph{propagation}. The latter is achieved by  
removing inconsistent values 
from the domains of variables. That way, values that cannot appear in a 
solution, given the current set of decisions, are never tried by the search. 
\emph{Global constraints} are higher order constraints that enforce a 
complex 
constraint over a set of variables. These are typically implemented with 
specialized algorithms that perform substantially better than decomposition approaches
\cite{regin1994filtering,beldiceanu2012global}.

@@

Note how this CP model is easy to adapt to more classic settings where no 
patterns are used (by dropping Eqs. \ref{EQ:patt} and \ref{EQ:pcost}), or 
without \emph{unknown} nodes (by dropping Eq. \ref{EQ:unk}).
\section{Results \label{SEC:Res}}

@@

\begin{figure}[!htb]
\begin{minipage}[b]{.5\linewidth}
\centering
\hspace*{-15pt}
\begin{tikzpicture}[baseline,scale=0.47]
\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    ylabel={},
		    xmin=1, xmax=28,
		    ymin=0.2, ymax=0.9,
		    xtick={1,5,10,15,20,25,28},
		    ytick={0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    legend pos=south east,
		    ymajorgrids=true,
		    grid style=dashed,
		    style = {font=\huge}
		]

@@

		\end{axis}
\end{tikzpicture}
\subcaption{Precision (museum)}\label{fig:museumprec}
\end{minipage}%
\begin{minipage}[b]{.5\linewidth}
\centering
\begin{tikzpicture}[baseline,scale=0.47]
		\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    %ylabel={recall},
		    xmin=1, xmax=28,
		    ymin=0.2, ymax=0.9,
		    xtick={1,5,10,15,20,25,28},
		    ytick={0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    ymajorgrids=true,
		    grid style=dashed,
		    		    style = {font=\huge},
		    legend style={at={(0.5,1.1)},anchor=south, legend columns=-1}
		]

@@

\begin{minipage}[b]{.5\linewidth}
\centering
\hspace*{-15pt}
\begin{tikzpicture}[baseline,scale=0.47]
\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    ylabel={},
		    xmin=1, xmax=11,
		    ymin=0.5, ymax=0.9,
		    xtick={1,2,4,6,8,10,11},
		    ytick={0.5,0.6,0.7,0.8,0.9,1},
		    legend pos=south east,
		    ymajorgrids=true,
		    grid style=dashed,
		    		    		    style = {font=\huge},
		]

@@

		\end{axis}
\end{tikzpicture}
\subcaption{Precision (soccer)}\label{fig:soccerprec}
\end{minipage}%
\begin{minipage}[b]{0.5\linewidth}
\centering
\begin{tikzpicture}[baseline,scale=0.47]
		\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    %ylabel={recall},
		    xmin=1, xmax=11,
		    ymin=0.3, ymax=0.9,
		    xtick={1,2,4,6,8,10,11},
		    ytick={0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    ymajorgrids=true,
		    grid style=dashed, legend pos=south east,
		    		    		    style = {font=\huge}
		]

@@

\begin{table}[!ht]\small
  \centering
  \caption{Example table.}
  	\label{tab:semlab}
  	\begin{tabular}{ccc} 
  		\hline
		Person & Age & Gender \\
  		\hline
	    John & 20 & M\\
    	Jane & 30 & F\\
  		\hline
		\end{tabular} 
%\vspace*{-3mm}
\end{table}

\begin{table}[!ht]
	\centering
	\caption{Feature extractors that produce a scalar value}
	\begin{tabular}{p{3cm}p{4.5cm}}
	Feature name & Description \\ \hline
	num-unique-vals & A simple count of unique entries of a column\\
	\rowcolor{Gray}
	prop-unique-vals & The proportion of unique entries of a column. \\
	\rowcolor{Gray}
	prop-missing-vals & The proportion of all missing/empty entries from all entries in a column.\\
	ratio-alpha-chars & The average of the proportion of alphabetic characters from each entry.\\
	\rowcolor{Gray}
	prop-numerical-chars & The average of the proportion of numeric characters from each entry.\\
	prop-whitespace-chars    & The average of the proportion of whitespace characters from each entry.\\
	\rowcolor{Gray}
	prop-entries-with-at-sign& The proportion of entries with an '@' sign. \\
	prop-entries-with-hyphen & The proportion of entries with a '-'. \\
	\rowcolor{Gray}
	prop-range-format        & The proportion of entries which follow some \\ numerical range format &  ([0-9]+)-([0-9]+) \\
	\rowcolor{Gray}
	is-discrete              & A binary indicator if the column entries are discrete. \\
	entropy-for-discrete-values & The entropy of the entries of a column.\\
	\rowcolor{Gray}	
	shannon-entropy          & Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'\\
	\end{tabular}
\end{table}
\ignore{
The following feature extractors produce a scalar value:
\begin{enumerate}
\item num-unique-vals      : A simple count of unique entries of a column
\item prop-unique-vals     : The proportion of unique entries of a column. 
\item prop-missing-vals    : The proportion of all missing/empty entries from all entries in a column.
\item ratio-alpha-chars    : The average of the proportion of alphabetic characters from each entry.
\item prop-numerical-chars : The average of the proportion of numeric characters from each entry.
\item prop-whitespace-chars       : The average of the proportion of whitespace characters from each entry.
\item prop-entries-with-at-sign   : The proportion of entries with an '@' sign.
\item prop-entries-with-hyphen    : The proportion of entries with a '-'.
\item prop-range-format           : The proportion of entries which follow some numerical range format: ([0-9]+)-([0-9]+).
\item is-discrete                 : A binary indicator if the column entries are discrete.
\item entropy-for-discrete-values : The entropy of the entries of a column.
\item shannon-entropy             : Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'
\end{enumerate}
}

\begin{table}[!ht]
	\centering
	\caption{Feature extractors that produce a vector value. Some of these are computed together to reduce computation time.
	Others are values that are computed for all classes.}

@@

\ignore{
The following feature extractors produce a vector value.
Some of these are computed together to reduce computation time.
Others are values that are computed for all classes.

\begin{enumerate}
\item inferred-data-type      : Produces a vector which contains the inferred data type.  
                          Only a single element is non-zero in the vector (i.e. 1-of-K format).  
                          The set of inferred types are in {float,integer,long,boolean,date,time,datetime,string}.

\item char-dist-features      : produces a vector for character distribution of column values;
                        considered characters are printable characters:
%                        0123456789abcdefghijklmnopqrstuvwxyz
%                        ABCDEFGHIJKLMNOPQRSTUVWXYZ!"
%                        #$%&\'()*+,-./:;<=>?@[\\]^_`{|}~ \t\n\r\x0b\x0c

\item stats-of-text-length    : Computes the mean, median, mode, min, max of string lengths from a columns entries.

\item stats-of-numerical-type : Computes the mean, median, mode, min, max of entries from a column of numerical type.

\item prop-instances-per-class-in-knearestneighbours         : This finds the k columns from the training set with most similar names.
                                                         It then produces a vector of class proportions from these k neighbours.
                                                         Parameters:
                                                             "num-neighbours": The number of neighbours to find (k).

\item mean-character-cosine-similarity-from-class-examples   : For some test instance t, this computes the character distribution 
                                                         from this instance's entries.  It then produces a vector of the average
                                                         cosine similarity of t with training instances of each class.

\item min-editdistance-from-class-examples                   : For some test instance t, this produces a vector of the minimum column name
                                                         edit distance between t and all training instances of each class.
                                                         Parameters:
                                                              "cache-size": Size of cache for storing edit distances of words 
                                                              "max-comparisons-per-class": Max number of training instances to compare
                                                                                           against for each class.

\item min-wordnet-jcn-distance-from-class-examples            : For some test instance t, this produces a vector of the minimum column name
                                                         JCN WS4J distance between t and all training instances of each class.
                                                         Parameters:
                                                              "cache-size": Size of cache for storing wordnet distances of words 
                                                              "max-comparisons-per-class": Max number of training instances to compare
                                                                                           against for each class.

\item min-wordnet-lin-distance-from-class-examples            : For some test instance t, this produces a vector of the minimum column name
                                                         LIN WS4J distance between t and all training instances of each class.
                                                         Parameters:
                                                              "cache-size": Size of cache for storing wordnet distances of words 
                                                              "max-comparisons-per-class": Max number of training instances to compare
                                                                                           against for each class.

\end{enumerate}
}

\subsection{\minizinc{} Model \label{Ann:MZ}}

>>>>

\usepackage{colortbl}
\usepackage{color}
\usepackage{xcolor}
\definecolor{Gray}{gray}{0.96}
\usepackage{amsmath,amsthm,stmaryrd}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algrenewcommand\algorithmiccomment[2][\footnotesize]{{#1\hfill\(\triangleright\)
 #2}} %\normalsize
\usepackage{amssymb}
\usepackage{textcomp}

@@

\usepackage{listings}
%%AR defines some stuff such as syntax highlighting for stochastic Minzinc listings
\input{commondefs.tex}

@@

\subsubsection{Example} Consider the situation where we have some simple relational database tables with columns $\langle$\textit{Surname, Workplace, Date}$\rangle$, and 
$\langle$\textit{Company, Event, Begin, Until, Location}$\rangle$.
We do not \textit{a priori} know if ``Date'' in the first table is a 
date of birth or the date a person started working; or whether the 
``Location'' on the second table refers to the location of the event, or the 
address of the company. 
Given an ontology like the one in Fig. \ref{FIG:onto}, we would like to automatically map the new data sources to classes of the 
ontology.\hfill$\square$

@@

In our work we consider that an ontology $\mathcal{O}$ includes basic elements 
as classes which represent a set of concepts, literal values, individuals which 
are members of classes and properties~\cite{Spanos:semweb}.
Properties are classified into \emph{object properties}, which relate two 
individuals, 
and \emph{datatype properties}, which relate individuals to literal values.
Fig.~\ref{FIG:onto} gives an example of an ontology. 
Here, an individual of the class ``Organisation'' can be related to an individual of the class ``Person'' via object properties ``ceo'' or ``worksFor'', and it can have a data property ``name''. 
We also consider a special ``subclass'' (or ``is-a'') type of object properties.
For 
example, ``Cities'' and ``States'' are both ``Places''. 
A subclass inherits all the properties of the parent class.

@@

The algorithm we use for the construction of the alignment graph 
$\mathcal{G_O}$ was given by~\citeasnoun{taheriyan2016learning}.
Briefly, it has three parts:
%\begin{enumerate}
$(i)$ %\item 
Adding the known semantic models
$(ii)$ %\item 
Adding the semantic types learned for the target source
%\item 
$(iii)$ Expanding the graph using the domain ontology.
%\end{enumerate}

@@

The graph is weighted by a function $w_\mathcal{O} : \mathcal{E_O} \mapsto 
\mathbb{R}$ such that edges which are present in
the known semantic models have lower weights than those 
which are inferred from the ontology.
In the next section we will see that this makes edges from semantic models 
more attractive. 
\citeasnoun{taheriyan2016learning} provide details on the weighting function.
%\ddg{How exactly are they weighted?}

@@

A Constraint Optimization Problem (COP) is a tuple $P=(\boldsymbol{v},C,o)$ 
where $\boldsymbol{v}$ is a 
set of \emph{variables}, $C$ is a set of $n$-ary 
\emph{constraints} over variables $\boldsymbol{v}$ and $o$ is an 
\emph{objective function} $o : \boldsymbol{v} \mapsto \mathbb{R}$ to be 
(w.l.o.g.) minimized. 
A valuation such that all variables in $\boldsymbol{v}$ map 
to exactly one value, and all the constraints in $C$ are satisfied by the 
valuation is a \emph{solution} to $P$. A solution $\theta$ is optimal if$\not 
\exists \theta', \theta'(\boldsymbol{v}) < \theta(\boldsymbol{v})$.

@@

Briefly, the way a CP solver works is by assigning a value to each variable of 
$\boldsymbol{v}$ in turn. 
\ignore{At each iteration it will check whether any 
constraint in $C$ is violated, in which case it will backtrack to change its 
last decision. \ignore{The choice of the order of the variables to be assigned, 
and the 
choice of the value to be assigned to each variable is called the \emph{search 
strategy}.}}
CP uses a combination of complete search and \emph{propagation}. The latter
removes inconsistent values 
from the domains of variables. That way, values that cannot appear in a 
solution, given the current set of decisions, are never tried by the search. 
\emph{Global constraints} are higher order constraints that enforce a 
complex 
constraint over a set of variables. They are implemented with 
specialized algorithms for performance.

@@

Note how this CP model is easy to adapt to more classic settings where no 
patterns are used (by dropping Eqs. \ref{EQ:patt} and \ref{EQ:pcost}), or 
without \emph{unknown} nodes (by dropping Eq. \ref{EQ:unk}).

We choose a search strategy that will first try to match attributes using the cheapest edges in $\mathcal{M}_\mathcal{O}^{s^\star}$. Then, it will try to use the cheapest unfixed edge $e\in\mathcal{E_O}$ that completes a pattern $p$ such that $w_\mathcal{P}(p) > w_\mathcal{O}(e)$ or simply the cheapest edge if none complete a pattern.

\section{Results \label{SEC:Res}}

@@

\begin{figure}[!htb]
\begin{minipage}[b]{.5\linewidth}
\centering
\hspace*{-15pt}
\begin{tikzpicture}[baseline,scale=0.43]
\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    ylabel={},
		    xmin=1, xmax=28,
		    ymin=0.2, ymax=0.9,
		    xtick={1,5,10,15,20,25,28},
		    ytick={0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    legend pos=south east,
		    ymajorgrids=true,
		    grid style=dashed,
		    style = {font=\huge}
		]

@@

		\end{axis}
\end{tikzpicture}
\subcaption{Precision (museum)}\label{fig:museumprec}
\end{minipage}%
\begin{minipage}[b]{.5\linewidth}
\centering
\begin{tikzpicture}[baseline,scale=0.43]
		\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    %ylabel={recall},
		    xmin=1, xmax=28,
		    ymin=0.2, ymax=0.9,
		    xtick={1,5,10,15,20,25,28},
		    ytick={0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    ymajorgrids=true,
		    grid style=dashed,
		    		    style = {font=\huge},
		    legend style={at={(0.5,1.1)},anchor=south, legend columns=-1}
		]

@@

\begin{minipage}[b]{.5\linewidth}
\centering
\hspace*{-15pt}
\begin{tikzpicture}[baseline,scale=0.43]
\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    ylabel={},
		    xmin=1, xmax=11,
		    ymin=0.5, ymax=0.9,
		    xtick={1,2,4,6,8,10,11},
		    ytick={0.5,0.6,0.7,0.8,0.9,1},
		    legend pos=south east,
		    ymajorgrids=true,
		    grid style=dashed,
		    		    		    style = {font=\huge},
		]

@@

		\end{axis}
\end{tikzpicture}
\subcaption{Precision (soccer)}\label{fig:soccerprec}
\end{minipage}%
\begin{minipage}[b]{0.5\linewidth}
\centering
\begin{tikzpicture}[baseline,scale=0.43]
		\begin{axis}[
		    xlabel={\emph{known semantic models}},
		    %ylabel={recall},
		    xmin=1, xmax=11,
		    ymin=0.3, ymax=0.9,
		    xtick={1,2,4,6,8,10,11},
		    ytick={0.3,0.4,0.5,0.6,0.7,0.8,0.9,1},
		    ymajorgrids=true,
		    grid style=dashed, legend pos=south east,
		    		    		    style = {font=\huge}
		]

@@

\begin{table}[!ht]\small
  \centering
  \caption{Example table.}
  	\begin{tabular}{ccc} 
  		\hline
		Person & Age & Gender \\
  		\hline
	    John & 20 & M\\
    	Jane & 30 & F\\
  		\hline
		\end{tabular} 
%\vspace*{-3mm}
\end{table}

\begin{table}[h]
	\centering
	\caption{Feature extractors that produce a scalar value}
	\begin{tabular}{p{3cm}p{4.5cm}}
	Feature name & Description \\ \hline
	num-unique-vals & A simple count of unique entries of a column\\
	\rowcolor{Gray}
	prop-unique-vals & The proportion of unique entries of a column. \\
	\rowcolor{Gray}
	prop-missing-vals & The proportion of all missing/empty entries from all entries in a column.\\
	ratio-alpha-chars & The average of the proportion of alphabetic characters from each entry.\\
	\rowcolor{Gray}
	prop-numerical-chars & The average of the proportion of numeric characters from each entry.\\
	prop-whitespace-chars    & The average of the proportion of whitespace characters from each entry.\\
	\rowcolor{Gray}
	prop-entries-with-at-sign& The proportion of entries with an '@' sign. \\
	prop-entries-with-hyphen & The proportion of entries with a '-'. \\
	\rowcolor{Gray}
	prop-range-format        & The proportion of entries which follow some \\ numerical range format &  ([0-9]+)-([0-9]+) \\
	\rowcolor{Gray}
	is-discrete              & A binary indicator if the column entries are discrete. \\
	entropy-for-discrete-values & The entropy of the entries of a column.\\
	\rowcolor{Gray}	
	shannon-entropy          & Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'\\
	\end{tabular}
\end{table}

\begin{table}[t]
	\centering
	\caption{Feature extractors that produce a vector value. Some of these are computed together to reduce computation time.
	Others are values that are computed for all classes.}

@@

\clearpage\newpage
\subsection{\minizinc{} Model \label{Ann:MZ}}
\renewcommand{\lstlistingname}{Model}
\begin{minipage}{\textwidth}
%\begin{figure}[!h]
	\begin{lstlisting}[caption={\minizinc{} model (part 1 of 2)},captionpos=b]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% MiniZinc model to solve the Steiner Tree Problem arising %%
%% in the Rel2Onto problem.                                 %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

include "globals.mzn";

%%%%%%% Steiner Tree Problem
int: nbV;                               % Nb of nodes in alignment graph
int: nbE;                               % Nb of edges in alignment graph
set of int: nodes = 1..nbV;    
set of int: edges = 1..nbE;
set of int: cnodes;                     % Class nodes in alignment graph
set of int: dnodes;                     % Data nodes in alignment graph
set of int: anodes;                     % Attribute nodes in alignment graph (normally none)
array [nodes] of var bool: vs;          % Is a node of the alignment graph in the solution?
array [edges] of var bool: es;          % Is an edge of the alignment graph in the solution?
array [nodes,edges] of bool: adjacent;  % adjacent[n,e] <=> node n is an endnode of edge e
array [edges] of nodes: heads;          % endpoint 1 of each edge (The graph is undirected!!)
array [edges] of nodes: tails;          % endpoint 2 of each edge
array[1..2,edges] of int: endpairs = array2d(1..2,edges,tails++heads);
array [edges] of int: ws;               % Weights of edges in alignment graph
var 0..sum(ws): w;                      % Weight of the solution corresponding to the alignment part
constraint steiner_tree(vs,es,adjacent,endpairs,w,ws);  % Steiner Tree constraint
constraint forall (d in dnodes) (vs[d] <-> sum(e in edges where adjacent[d,e])(es[e]) == 1); %Only 1 edges per data node

%%%%%%% Matching Problem
int: nbA;                                           % Nb of attribute nodes
set of int: atts = 1..nbA;                 
array [atts] of set of int: attribute_domains;      % Data nodes to which each attribute can attach  
constraint forall(att in anodes) (vs[att] == true); % Attribute nodes (normally none) are in the graph.
array [atts] of var int: match;                     % Match between Attributes and Data nodes
array[atts,int] of int: match_costs;                % Weight of match edges
constraint forall (a in atts) (match[a] in (attribute_domains[a]));   % Must match someone in the domain
constraint forall (a in atts) (vs[match[a]]);       % The data node matched must be in the subgraph of the alignment
constraint alldifferent(match);                     % Exact match
var 0..sum(a in atts)(sum(n in index_set_2of2(match_costs))(match_costs[a,n])): wm;  %Total cost from matching part :
constraint wm = sum(a in atts) (match_costs[a,match[a]]);

	\end{lstlisting}
%	\caption{\minizinc{} model (1/2).}
%\end{figure}
\end{minipage}
\clearpage
\newpage
\begin{minipage}{\textwidth}
%\begin{figure*}[t]
	\begin{lstlisting}[caption={\minizinc{} model (part 2 of 2)},captionpos=b]
%%%%%%% Unknown nodes
int: nb_unknown_nodes;                 % Number of unknown data nodes
set of int: unknowns = 1..nb_unknown_nodes;
array [int] of int: all_unk_nodes;     % List of all unknown nodes (first two elements are class nodes, by convention)
int: node_root = all_unk_nodes[1];     % Root node if unknown class node is used.
int: node_unk = all_unk_nodes[2];      % Unknown class node
constraint if nb_unknown_nodes > 0 then
    vs[node_unk] <-> vs[node_root]     % If class node "Unknown" is used, class node "Root" is used as well
 /\ vs[node_root] <-> sum(e in edges where adjacent[node_root,e])(es[e]) in {1,2}
else true %Ignore constraint
endif;

%%%%%%% Patterns
int: nb_patterns;                     %Number of patterns
set of int: PATTS = 1..nb_patterns;
array[PATTS] of int: patterns_w;      %Support of patterns
array[PATTS] of set of int: patterns; %The actual patterns
array[PATTS] of var bool: c_p;        %Is pattern chosen
constraint forall (p in PATTS) ((sum (e in patterns[p])   % All edges of a pattern are used <-> pattern is used
							                  	(es[e]) == card(patterns[p])) <-> c_p[p]);
var 0..sum(patterns_w): wp; % Total prize collected from patterns : 
constraint if nb_patterns = 0 then wp = 0 else wp = sum(p in PATTS) (c_p[p] * patterns_w[p]) endif;

%%%%%%% Final objective function 
var int: obj = w + wm - wp; % Total cost of semantic model - prize collected from patterns

%%%%%%% Search startegy
array[atts,int] of int: match_costs_sorted;  % Match of attributes, sorted by cost.
ann: cheap_matches = int_search([ match[a] == match_costs_sorted[a,i] |a in atts, i in index_set_2of2(match_costs_sorted)], input_order,indomain_max,complete);

array[edges] of int: sorted_edges = arg_sort(ws);
ann: cheap_edges = int_search([es[sorted_edges[nbE + 1 - e]] | e in edges],input_order,indomain_min,complete);

ann: cheap_edges_fill_pattern; % Custom search strategy for patterns
ann: custom(ann, array[int] of int, array[int] of var bool);
array[int] of ann: cefp = 
if nb_patterns == 0 then [] 
else [custom(cheap_edges_fill_pattern,[nbE]++[nb_patterns]++ws++[ card(patterns[p]) | p in PATTS]++patterns_w++array1d([ e | p in PATTS, e in patterns[p]]),es)] endif;

solve ::seq_search([cheap_matches]++cefp++[cheap_edges]) 
minimize obj;

	\end{lstlisting}
	%\caption{\minizinc{} model (2/2).}
%\end{figure*}
\end{minipage}
